---
layout: post
title: 'Human Resource Analytics: Data Exploration'
date: 2017-01-26
author: Yin-Ting 
categories: [Project]
tags: [R, Classification]
---
This project is based on the [dataset](https://www.kaggle.com/ludobenistant/hr-analytics) found in Kaggle. This dataset have employees' information including Employee satisfaction level, Last evaluation, Number of projects, Average monthly hours, Time spent at the company, Whether they have had a work accident, Whether they have had a promotion in the last 5 years, Department, Salary, and Whether the employee has left. The goal of this project is to successfully predict which kind of employees are highly possible to leave in the future. 

## Data Cleaning
```{r}
# Import Raw Data
dat=read.csv("/Users/chou/Google Drive/websites/github/myblog-master/data_project/HR_analytics.csv", header=TRUE)
data=dat # keep raw data pure

# Rename the variables
names(data)=c("satisf_level","last_eval","num_proj",
              "ave_mon_hrs","time_spend","work_accid",
              "left_or_not","promo_last_5yrs","department",
              "salary")

# Reorder the data
data=data[,c(1:6,8:10,7)]

# Correct variables' attributes
data$work_accid=as.factor(data$work_accid)
data$promo_last_5yrs=as.factor(data$promo_last_5yrs)
data$left_or_not=as.factor(data$left_or_not)
str(data) 

# Finding NA
which(is.na(data), arr.ind=TRUE) #the indices of NA values
```

***

## Data Exploration (Quantitative Explanotary Variables vs. Qualitative Binary Response Variable)
### R functions 
#### Table
```{r message = FALSE}
t_bygroup=function(d, x, y, n){
  # an elegant way to install a missing package
  if (!require("plyr")) install.packages("plyr")
    t <- ddply(d, y, .fun = function(dd){
               c(Mean = mean(dd[,x],na.rm=TRUE),
               Sd = sd(dd[,x],na.rm=TRUE),
               min=min(dd[,x]), 
               Q1=quantile(dd[,x],0.25),
               Median=quantile(dd[,x],0.5),
               Q3=quantile(dd[,x],0.75), 
               Max=max(dd[,x])) })
  return(t)
}
```
#### Histogram
```{r message = FALSE}
hist_bygroup=function(d,xx,yy,name){
    if (!require("ggplot2")) install.packages("ggplot2")
    ggplot(d, aes_string(x=xx, color=yy, fill=yy))+ 
    geom_histogram(aes(y=..density..), alpha=0.5, 
                 position="identity")+
    geom_density(alpha=.3)+
    ggtitle(name)  
}
```

#### Box plot
```{r message = FALSE}
box_bygroup=function(d,xx,yy,name){
    if (!require("ggplot2")) install.packages("ggplot2")
    ggplot(d, aes_string(x=yy, y=xx, fill=yy)) + 
    geom_boxplot()+
    ggtitle(name) 
}
```

### Employee satisfaction level
```{r message = FALSE, fig.width=10, fig.height=5}
# Table
t_bygroup(d=data, x="satisf_level", y='left_or_not', n=2)

if (!require("Rmisc")) install.packages("Rmisc") # Arrangement for multiple ggplots
# Histogram
hist_sat=hist_bygroup(d=data, xx="satisf_level", yy="left_or_not",
               "Histogram for satisf_level")
# Box plot
box_sat=box_bygroup(d=data, xx="satisf_level", yy="left_or_not",
               "Boxplot for satisf_level")
multiplot(plotlist = list(hist_sat,box_sat), cols = 2) 
```

### Last evaluation
```{r fig.width=10, fig.height=5}
# Table
t_bygroup(data, "last_eval", "left_or_not")

# Histogram
hist_eval=hist_bygroup(d=data, xx="last_eval", yy="left_or_not",
               "Histogram for last_eval")
# Box plot
box_eval=box_bygroup(d=data, xx="last_eval", yy="left_or_not",
               "Boxplot for last_eval")
multiplot(plotlist = list(hist_eval,box_eval), cols = 2) 

```

### Number of projects
```{r fig.width=10, fig.height=5}
# Table
t_bygroup(data, "num_proj", "left_or_not")
# Histogram
hist_proj=hist_bygroup(d=data, xx="num_proj", yy="left_or_not",
               "Histogram for num_proj")
# Box plot
box_proj=box_bygroup(d=data, xx="num_proj", yy="left_or_not",
               "Boxplot for num_proj")
multiplot(plotlist = list(hist_proj,box_proj), cols = 2) 

```

### Average monthly hours
```{r fig.width=10, fig.height=5}
# Table
t_bygroup(data, "ave_mon_hrs", "left_or_not")
# Histogram
hist_hrs=hist_bygroup(d=data, xx="ave_mon_hrs", 
                       yy="left_or_not",
                       "Histogram for ave_mon_hrs")
# Box plot
box_hrs=box_bygroup(d=data, xx="ave_mon_hrs",
                     yy="left_or_not",
                     "Boxplot for ave_mon_hrs")
multiplot(plotlist = list(hist_hrs,box_hrs), cols = 2) 
```

### Time spent at the company
```{r fig.width=10, fig.height=5}
# Table
t_bygroup(data, "time_spend", "left_or_not")
# Histogram
hist_time=hist_bygroup(d=data, xx="time_spend", 
                       yy="left_or_not",
                       "Histogram for time_spend")
# Box plot
box_time=box_bygroup(d=data, xx="time_spend",
                     yy="left_or_not",
                     "Boxplot for time_spend")
multiplot(plotlist = list(hist_time,box_time), cols = 2) 
```

***

## Data Exploration (Qualitative Explanotary Variables vs. Qualitative Binary Response Variable)
### R function

### Whether they have had a work accident 
```{r}
xtabs(~work_accid+left_or_not, data)
```

### Whether they have had a promotion in the last 5 years
```{r}
xtabs(~promo_last_5yrs+left_or_not, data)
```
### Department
```{r}
xtabs(~department+left_or_not, data)
```
### Salary
```{r}
xtabs(~salary+left_or_not, data)
```
